\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}

\newtheorem{definition}{Definition}[section]
\newtheorem{property}{Property}[section]
\newtheorem{requirement}{Requirement}[section]

\title{Brahim Secure Intelligence (BSI)\\
\large Formal System Specification v1.0}

\author{Elias Oulad Brahim}
\date{January 24, 2026}

\begin{document}

\maketitle

\begin{abstract}
This document provides the formal specification for the Brahim Secure Intelligence (BSI) system, a unified framework for cryptographic security and AI safety grounded in the mathematical structure of the golden ratio. All system parameters derive from the Brahim Security Constant $\beta = \sqrt{5} - 2 = 1/\varphi^3$.
\end{abstract}

\tableofcontents

\newpage

\section{Introduction}

\subsection{Purpose}

This specification defines the mathematical foundations, algorithms, and interfaces for the Brahim Secure Intelligence (BSI) system. BSI provides:

\begin{enumerate}
    \item Cryptographic primitives based on the Perfect Wormhole Transform
    \item AI safety verification via the ASIOS energy functional
    \item Intent routing through Brahim Sequence territory classification
    \item Cognitive architecture with 12-wavelength processing
\end{enumerate}

\subsection{Scope}

This specification covers:
\begin{itemize}
    \item Mathematical constants and their derivations
    \item Algorithm specifications with formal properties
    \item Interface definitions for system modules
    \item Security and safety requirements
\end{itemize}

\subsection{Notation}

\begin{table}[h]
\centering
\begin{tabular}{ll}
\toprule
\textbf{Symbol} & \textbf{Meaning} \\
\midrule
$\varphi$ & Golden ratio $(1+\sqrt{5})/2$ \\
$\beta$ & Brahim Security Constant $1/\varphi^3$ \\
$\alpha$ & Wormhole Attraction Constant $1/\varphi^2$ \\
$\mathcal{B}$ & Brahim Sequence \\
$S$ & Brahim Sum (214) \\
$C$ & Brahim Center (107) \\
$\bar{C}$ & Brahim Centroid $\mathcal{B}/S$ \\
$W^*$ & Perfect Wormhole Transform \\
$E[\psi]$ & Energy functional \\
$\mathcal{G}$ & Genesis Constant \\
\bottomrule
\end{tabular}
\caption{Notation Summary}
\end{table}

\section{Mathematical Foundations}

\subsection{Fundamental Constants}

\begin{definition}[Golden Ratio]
The golden ratio is defined as:
\begin{equation}
\varphi = \frac{1 + \sqrt{5}}{2} = 1.6180339887498949\ldots
\end{equation}
satisfying $\varphi^2 = \varphi + 1$.
\end{definition}

\begin{definition}[Brahim Security Constant]
The Brahim Security Constant is:
\begin{equation}
\beta = \frac{1}{\varphi^3} = \sqrt{5} - 2 = 2\varphi - 3 = 0.2360679774997897\ldots
\end{equation}
\end{definition}

\begin{property}[Polynomial Root]
$\beta$ satisfies $\beta^2 + 4\beta - 1 = 0$.
\end{property}

\begin{property}[Continued Fraction]
$\beta = [0; \overline{4}] = [0; 4, 4, 4, \ldots]$.
\end{property}

\begin{property}[Self-Similarity]
$\alpha / \beta = \varphi$ where $\alpha = 1/\varphi^2$.
\end{property}

\subsection{Golden Ratio Hierarchy}

\begin{definition}[Hierarchy]
The golden ratio hierarchy $\{H_n\}$ is defined by:
\begin{align}
H_0 &= \varphi \\
H_n &= H_{n-1} / \varphi = 1/\varphi^n \quad \text{for } n \geq 1
\end{align}
\end{definition}

\begin{table}[h]
\centering
\begin{tabular}{clll}
\toprule
$n$ & Name & Symbol & Value \\
\midrule
0 & Base & $\varphi$ & 1.618033988749895 \\
1 & Compression & $1/\varphi$ & 0.618033988749895 \\
2 & Attraction & $\alpha$ & 0.381966011250105 \\
3 & Security & $\beta$ & 0.236067977499790 \\
4 & Damping & $\gamma$ & 0.145898033750315 \\
\bottomrule
\end{tabular}
\caption{Golden Ratio Hierarchy}
\end{table}

\subsection{Brahim Sequence}

\begin{definition}[Brahim Sequence]
\begin{equation}
\mathcal{B} = (27, 42, 60, 75, 97, 121, 136, 154, 172, 187)
\end{equation}
\end{definition}

\begin{definition}[Derived Constants]
\begin{align}
S &= 214 \quad \text{(Sum constant)} \\
C &= S/2 = 107 \quad \text{(Center)} \\
D &= |\mathcal{B}| = 10 \quad \text{(Dimension)}
\end{align}
\end{definition}

\begin{property}[Critical Line]
$C/S = 1/2$, corresponding to the Riemann critical line $\Re(s) = 1/2$.
\end{property}

\section{Core Algorithms}

\subsection{Perfect Wormhole Transform}

\begin{algorithm}
\caption{Perfect Wormhole Transform $W^*$}
\begin{algorithmic}[1]
\Require $\sigma \in \mathbb{R}^D$ (identity signature)
\Ensure $W^*(\sigma) \in \mathbb{R}^D$ (transformed signature)
\State $\bar{C} \gets \mathcal{B} / S$ \Comment{Compute centroid}
\For{$i = 1$ to $D$}
    \State $W^*(\sigma)_i \gets \sigma_i / \varphi + \bar{C}_i \cdot \alpha$
\EndFor
\State \Return $W^*(\sigma)$
\end{algorithmic}
\end{algorithm}

\begin{property}[Fixed Point]
$W^*(\bar{C}) = \bar{C}$.
\end{property}

\begin{property}[Compression]
$\|W^*(\sigma)\| / \|\sigma\| \to 1/\varphi$ as $\sigma \to \infty$.
\end{property}

\subsection{Inverse Wormhole Transform}

\begin{algorithm}
\caption{Inverse Wormhole Transform $W^{*-1}$}
\begin{algorithmic}[1]
\Require $y \in \mathbb{R}^D$ (transformed signature)
\Ensure $\sigma \in \mathbb{R}^D$ (original signature)
\State $\bar{C} \gets \mathcal{B} / S$
\For{$i = 1$ to $D$}
    \State $\sigma_i \gets (y_i - \bar{C}_i \cdot \alpha) \cdot \varphi$
\EndFor
\State \Return $\sigma$
\end{algorithmic}
\end{algorithm}

\begin{property}[Invertibility]
$W^{*-1}(W^*(\sigma)) = \sigma$ for all $\sigma \in \mathbb{R}^D$.
\end{property}

\subsection{Energy Functional}

\begin{definition}[Density Function]
For vector $x \in \mathbb{R}^n$:
\begin{equation}
\text{density}(x) = \frac{\text{Var}(x)}{\text{Mean}(x)}
\end{equation}
\end{definition}

\begin{definition}[Energy Functional]
For state $\psi$ and Hamiltonian weights $w$:
\begin{equation}
E[\psi] = \left(\text{density}(\psi + w) - \mathcal{G}\right)^2
\end{equation}
where $\mathcal{G} = \beta / C$ is the Genesis Constant.
\end{definition}

\begin{algorithm}
\caption{Safety Assessment}
\begin{algorithmic}[1]
\Require $\psi \in \mathbb{R}^n$ (embedding state)
\Ensure Verdict $\in \{\text{SAFE}, \text{NOMINAL}, \text{CAUTION}, \text{UNSAFE}, \text{BLOCKED}\}$
\State $d \gets \text{density}(\psi)$
\State $E \gets (d - \mathcal{G})^2$
\If{$E < 10^{-6}$}
    \State \Return SAFE
\ElsIf{$E < 10^{-4}$}
    \State \Return NOMINAL
\ElsIf{$E < 10^{-2}$}
    \State \Return CAUTION
\ElsIf{$E < 0.1$}
    \State \Return UNSAFE
\Else
    \State \Return BLOCKED
\EndIf
\end{algorithmic}
\end{algorithm}

\section{System Modules}

\subsection{Module: BrahimConstants}

\begin{requirement}[R-CONST-01]
The module SHALL provide all fundamental constants with IEEE 754 double precision.
\end{requirement}

\begin{requirement}[R-CONST-02]
The module SHALL provide verification functions for all mathematical identities.
\end{requirement}

\textbf{Interface:}
\begin{verbatim}
object BrahimConstants {
    val PHI: Double
    val BETA_SECURITY: Double
    val ALPHA_WORMHOLE: Double
    val BRAHIM_SEQUENCE: IntArray
    val BRAHIM_SUM: Int
    val BRAHIM_CENTER: Int

    fun verifyBetaIdentities(): Map<String, Boolean>
    fun verifyHierarchy(): Map<String, Boolean>
    fun getCentroid(): DoubleArray
}
\end{verbatim}

\subsection{Module: WormholeCipher}

\begin{requirement}[R-CIPHER-01]
The cipher SHALL derive S-box from $\beta$ continued fraction.
\end{requirement}

\begin{requirement}[R-CIPHER-02]
The cipher SHALL use key-derived secret centroid, not public $\bar{C}$.
\end{requirement}

\begin{requirement}[R-CIPHER-03]
The cipher SHALL use nonce-based construction with 128-bit nonces.
\end{requirement}

\textbf{Interface:}
\begin{verbatim}
class WormholeCipher(masterKey: ByteArray) {
    fun encrypt(plaintext: ByteArray): ByteArray
    fun decrypt(ciphertext: ByteArray): ByteArray
    fun wormholeTransform(sigma: DoubleArray): WormholeResult
    fun inverseWormhole(y: DoubleArray): DoubleArray
}
\end{verbatim}

\subsection{Module: ASIOSGuard}

\begin{requirement}[R-SAFETY-01]
The guard SHALL compute energy functional for all input states.
\end{requirement}

\begin{requirement}[R-SAFETY-02]
The guard SHALL block states with $E[\psi] \geq 0.1$.
\end{requirement}

\begin{requirement}[R-SAFETY-03]
The guard SHALL derive thresholds from $\beta$, not empirical values.
\end{requirement}

\textbf{Interface:}
\begin{verbatim}
class ASIOSGuard {
    fun computeEnergy(embedding: DoubleArray): Double
    fun assessSafety(embedding: DoubleArray): SafetyAssessment
    fun computeCorrectionGradient(embedding: DoubleArray,
                                   weights: DoubleArray): DoubleArray
}
\end{verbatim}

\subsection{Module: IntentRouter}

\begin{requirement}[R-ROUTER-01]
The router SHALL classify queries into exactly 10 territories.
\end{requirement}

\begin{requirement}[R-ROUTER-02]
Territory indices SHALL correspond to Brahim Sequence positions.
\end{requirement}

\begin{requirement}[R-ROUTER-03]
The router SHALL apply wormhole transform before distance computation.
\end{requirement}

\textbf{Interface:}
\begin{verbatim}
class IntentRouter {
    fun route(query: String): RoutingResult
    fun routeWithAnalysis(query: String): Map<String, Any>
    fun textToEmbedding(text: String): DoubleArray
}

enum class Territory {
    GENERAL, MATH, CODE, SCIENCE, CREATIVE,
    ANALYSIS, SYSTEM, SECURITY, DATA, META
}
\end{verbatim}

\subsection{Module: BOAAgent}

\begin{requirement}[R-AGENT-01]
The agent SHALL implement 12-wavelength cognitive architecture.
\end{requirement}

\begin{requirement}[R-AGENT-02]
The agent SHALL invoke ASIOSGuard at wavelength 6 (epsilon).
\end{requirement}

\begin{requirement}[R-AGENT-03]
The agent SHALL use $\beta$ as default learning rate.
\end{requirement}

\textbf{Interface:}
\begin{verbatim}
class BOAAgent {
    fun process(message: String): AgentResponse
    fun getAgentInfo(): Map<String, Any>
    fun reset()
}

data class AgentResponse(
    val content: String,
    val territory: Territory,
    val confidence: Double,
    val safetyScore: Double,
    val wavelengthStates: Map<String, Double>,
    val processingTime: Long
)
\end{verbatim}

\section{Security Requirements}

\begin{requirement}[R-SEC-01]
Master keys SHALL be at least 256 bits.
\end{requirement}

\begin{requirement}[R-SEC-02]
Nonces SHALL be unique per encryption operation.
\end{requirement}

\begin{requirement}[R-SEC-03]
S-box SHALL provide non-linearity with differential probability $\leq 2^{-6}$.
\end{requirement}

\begin{requirement}[R-SEC-04]
Key-derived centroids SHALL be computationally indistinguishable from random.
\end{requirement}

\section{Safety Requirements}

\begin{requirement}[R-SAFE-01]
All AI outputs SHALL pass through ASIOSGuard before delivery.
\end{requirement}

\begin{requirement}[R-SAFE-02]
States with verdict BLOCKED SHALL NOT produce output.
\end{requirement}

\begin{requirement}[R-SAFE-03]
Energy threshold derivations SHALL be traceable to $\beta$.
\end{requirement}

\section{Validation}

\begin{requirement}[R-VAL-01]
System SHALL verify all $\beta$ identities on startup.
\end{requirement}

\begin{requirement}[R-VAL-02]
System SHALL log verification results.
\end{requirement}

\begin{requirement}[R-VAL-03]
System SHALL refuse to start if verification fails.
\end{requirement}

\section{Appendix: Test Vectors}

\subsection{Constant Verification}
\begin{verbatim}
PHI = 1.6180339887498949
BETA = 0.2360679774997897
ALPHA = 0.3819660112501051

beta^2 + 4*beta - 1 = 4.44e-16 (accept if < 1e-14)
alpha/beta = 1.618033988749895 (must equal PHI)
C/S = 0.5 (exact)
\end{verbatim}

\subsection{Wormhole Transform}
\begin{verbatim}
Input:  sigma = [0.1, 0.2, 0.3, ..., 1.0]
Output: W*(sigma) = [0.110, 0.199, 0.293, ...]
Compression ratio: 0.618 (within 0.01 of 1/PHI)
\end{verbatim}

\subsection{Energy Functional}
\begin{verbatim}
Target: GENESIS_CONSTANT = 0.00221888
Safe state: density = 0.00221, E = 7.8e-11 -> SAFE
Unsafe state: density = 0.5, E = 0.248 -> BLOCKED
\end{verbatim}

\end{document}
